---
title: "Assignment_1"
author: "Melissa Jaffe & Dylan White"
date: "9/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
path <- "C:/Users/Melissa/Documents/Ecostats"
setwd(path)
knitr::opts_knit$set(root.dir = path)
```

```{r}
library(lubridate)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(scales)
library(tidyverse)
```

```{r}
clim_tree_data <- read.csv("tahoe_snotel_data.csv", header = T)
na.omit(clim_tree_data)  #first 10 days of 1995 ommitted
head(clim_tree_data)
sapply(clim_tree_data,class)
summary(clim_tree_data)
str(clim_tree_data)
is.data.frame(clim_tree_data)
```
#Question 1: Summarize the data by monthly normals (means) for each site
```{r}
subset_clim <- clim_tree_data %>% 
  group_by(site, month, year) %>%
  summarize(mean(tavg, na.rm = TRUE))
subset_clim

subset_clim_2 <- clim_tree_data %>% 
  group_by(site, elev, month, year) %>%
  summarize(mean(tavg, na.rm = TRUE))
subset_clim_2

head(subset_clim_2)
summary(subset_clim)


names(subset_clim)[names(subset_clim) == "mean(tavg, na.rm = TRUE)"] <- "t_average"
names(subset_clim_2)[names(subset_clim_2) == "mean(tavg, na.rm = TRUE)"] <- "t_average"


subset_clim_3 <- clim_tree_data %>% 
  group_by(site, month) %>% 
  summarise(mean(tavg,na.rm=T))

head(subset_clim_3)
summary(subset_clim_3)

names(subset_clim_3)[names(subset_clim_3) == "mean(tavg, na.rm = T)"] <- "t_average"


subset_clim_4<- clim_tree_data %>%
  group_by(site,elev,month) %>%
  summarise(mean(tavg,na.rm=T))
subset_clim_4

names(subset_clim_4)[names(subset_clim_4) == "mean(tavg, na.rm = T)"] <- "t_average"

is.data.frame(subset_clim)
  
```
#Question 2: Produce univariate summary statistics for data. Produce univariate plots characterizing the data
```{r}
hist_clim <- ggplot(subset_clim, aes(t_average)) + geom_histogram() + facet_wrap(~site)
hist_clim

year_hist <- ggplot(subset_clim, aes(year)) + geom_histogram() + facet_wrap(~site)
year_hist #not all sites have all years

elev_dot <- ggplot(subset_clim_2, aes(x = site, y = elev)) + geom_point() 
elev_dot

hist_clim_3 <- ggplot(subset_clim_3, aes(t_average)) + geom_histogram() + facet_wrap(~site)
hist_clim_3

clim_3_plot <- ggplot(subset_clim_3, aes(month,t_average)) +
                  geom_point()+  facet_wrap(~site)+labs(title = "Tahoe SNOTEL sites monthly normals",x="Temperature", y="Month" )

clim_3_plot


boot_mean_3 <- numeric()
B<- 1000
for(i in 1:B){
  boot.sample<-sample(subset_clim_3$t_average,size=length(subset_clim_3$t_average), replace=T)
  boot_mean_3[i]<-mean(boot.sample)
}
hist(boot_mean_3)



```
#Question 3: Fit a linear model that relates the average temperature in May to elevation using data from the 16 sites. Do the same for average temperatures in December. What are the lapse rates from these months? What do these values mean? 
```{r}
new_data_1 <- clim_tree_data %>%
  group_by(site, month) %>%
  summarize(mean(elev, na.rm = TRUE), mean(tavg, na.rm = TRUE))
names(new_data_1)[names(new_data_1) == "mean(tavg, na.rm = TRUE)"] <- "t_average"
names(new_data_1)[names(new_data_1) == "mean(elev, na.rm = TRUE)"] <- "elev"
head(new_data_1)

may_data <- clim_tree_data %>%
 subset(clim_tree_data$month == "May") %>%
  group_by(site, elev) %>%
  summarize(mean(tavg, na.rm = TRUE))
names(may_data)[names(may_data) == "mean(tavg, na.rm = TRUE)"] <- "t_average"

dec_data <- clim_tree_data %>%
 subset(clim_tree_data$month == "December") %>%
  group_by(site, elev) %>%
  summarize(mean(tavg, na.rm = TRUE))
names(dec_data)[names(dec_data) == "mean(tavg, na.rm = TRUE)"] <- "t_average"

  
may_lm <- lm(may_data$elev~may_data$t_average)
summary(may_lm)

dec_lm <- lm(dec_data$elev~dec_data$t_average)
summary(dec_lm)

```
#Question 4: Plot the relationship
```{r}

may_plot <- ggplot(may_data, aes(x = elev, y= t_average)) + geom_point() +geom_smooth(method = lm)
may_plot

dec_plot <-  ggplot(dec_data, aes(x = elev, y= t_average)) + geom_point() +geom_smooth(method = lm)
dec_plot
#without ggplot
plot(may_data$elev, may_data$t_average, xlab="Elevation", ylab ="Temp Average")
abline(lm(may_data$elev~may_data$t_average), col = "red")

plot(dec_data$elev, dec_data$t_average, xlab="Elevation", ylab ="Temp Average")
abline(lm(dec_data$elev~dec_data$t_average), col = "red")

```
#Question 5: Produce diagnostic  plots characterizing the fit of the linear model and its assumptions
```{r}
may_resids <- resid(may_lm)
may_y_hat <- fitted(may_lm)
plot(may_y_hat, may_resids)
qnorm(may_resids)
plot(may_lm)

dec_resids <- resid(dec_lm)
dec_y_hat <- fitted(dec_lm)
plot(dec_y_hat, dec_resids)
qnorm(dec_resids)
plot(dec_lm)
```
#Extra Credit: Develop a R script that automatically fits the temperature to elevation relationship for each months data. This script should automatically save out the month name, the environmental lapse rate, and the r-squared value of the fit for each month into a matrix. Then plot the lapse rate and rsquared value by month. Discuss the temporal variability of these in your write up.
```{r}

lapserate.function<- function(l){
  for(i in 1:length(subset_clim_4$month) )#192 long?
    lr_lm<- lm(subset_clim_4$elev~subset_clim_4$t_average)# needs for each month
    
    
    return(matrix(subset_clim_4$name,lr_lm$coef[2],summary(lr_lm)$r.squared)
    return(
      ggplot(subset_clim_4, aes(month, )) + geom_point(lr_lm$coef[2]) +
        geompoint(summary(lr_lm)$r.squared)) #my ggplot skills sucks
}

#not final calculations (i.e. "play space")
lr_lm<- lm(subset_clim_4$elev~subset_clim_4$t_average)

summary(lr_lm)
plot(lr_lm)
lr_lm$coef[2]
summary(lr_lm)$r.squared
lapse_rate<-lr_lm$coef[2]
lapse_rate
ggplot(lr_lm$coef[2], aes(month))
+ geom_point() +
        geom_point(summary(lr_lm)$r.squared))
length(subset_clim_4$month)
plot(subset_clim_4$month,subset_clim_4$t_average)
 ggplot(subset_clim_4)+ aes(month, ) + geom_point(lr_lm$coef[2]) +
        geompoint(summary(lr_lm)$r.squared))


```
